---
title: "Lab4"
author: "Nehir"
date: "2024-09-26"
output: "github_document"
---

```{r}
library(car)
library(dplyr)
library(tidyr)
library(ggplot2)
library(psych)
library(bruceR)

```

# Load dataset
```{r}

lab4data <- read.csv("/Users/nehirarpat/Documents/GitHub/Lab4/skewed.csv")


```


# Normality
## Normality Plots
```{r}


ggplot(lab4data, aes(x = Performance)) + geom_histogram(binwidth = 20) + theme_classic()

ggplot(lab4data, aes(x = Performance)) + geom_density(adjust = 2)  + theme_classic()

qq<-ggplot(lab4data, aes(sample = Performance)) + geom_qq()  + theme_classic()

qq+ geom_qq_line()


#functions below are base R functions, they can do the job, but not as customizable as ggplot

hist(lab4data$Performance)

qqnorm(lab4data$Performance, col = "steelblue", lwd = 2)

plot(density(lab4data$Performance, na.rm = TRUE, bw = 90),  lwd=2, main = "")

```

### Normality Plots by Group
```{r}

#a fun function to plot the violin plot by group, but it's not part of ggplot
violinBy(Performance ~ Group, data = lab4data, rain= TRUE, vertical = FALSE)


#Find a way to plot the histograms, density, and qq plots by groups using ggplot

ggplot(lab4data, aes(x = Performance)) + geom_histogram(binwidth = 20) + facet_wrap(~Group) + theme_classic()

ggplot(lab4data, aes(x = Performance)) + geom_density(adjust = 2)  + facet_wrap(~Group) + theme_classic()

qq<-ggplot(lab4data, aes(sample = Performance)) + geom_qq()  + facet_wrap(~Group) + theme_classic()

qq+ geom_qq_line()


#Use ggplot to create a violin plot by groups


```


## Normality Tests
```{r}

describe(lab4data$Performance)

shapiro.test(lab4data$Performance)


```

### Normality Tests by Group
```{r}

#Use the describeBy() function to get skewness and kurtosis by group
?describeBy()
describeBy(Performance ~ Group, data = lab4data)

#Use the group by function to get shapiro test results by group
lab4data %>%
  group_by(Group) %>%
  summarize(W = shapiro.test(Performance)$statistic, p_value = shapiro.test(Performance)$p.value)

#Use the filter function to get both





```
# Q1: What is your overall conclusion about this variable's normality? Why?
The test result revealed non-significant results which indicates that the variable is not normally distributed.


# Equal Variance between Groups
## Descrptive Variance
```{r}
lab4data_clean <- drop_na(lab4data)
var(lab4data_clean$Performance)

lab4data_clean %>%
  group_by(Group) %>%
  summarize(variance = var(Performance))


```

## Equal Variance Test
```{r}


leveneTest(Performance~Group, lab4data)

MANOVA(lab4data, dv = "Performance", between = "Group")

#What if you want to test equal variance between 2 groups specifically? 

lab4dataConG1<-lab4data %>%
  filter(Group == "Control" | Group == "G1")

leveneTest(Performance~Group, lab4dataConG1)


```

# Q2: Overall, does it meet the equal variance assumption across the groups? Why?
No because the test reveals that the p value is significant (it has reverse meaning).
# Transformation
```{r}

#if any of the assumption is not met, use transformation 

lab4data$Performance_log <- log10(lab4data$Performance)

ggplot(lab4data, aes(x = Performance_log)) + geom_histogram(binwidth = .2) + theme_classic()

ggplot(lab4data, aes(x = Performance_log)) + geom_density(adjust = 2)  + theme_classic()

qq<-ggplot(lab4data, aes(sample = Performance_log)) + geom_qq()  + theme_classic()

qq+ geom_qq_line()


ggplot(lab4data, aes(x = Performance_log)) + geom_histogram(binwidth = .2) + facet_wrap(~Group) + theme_classic()

ggplot(lab4data, aes(x = Performance_log)) + geom_density(adjust = 2)  + facet_wrap(~Group) + theme_classic()

qq<-ggplot(lab4data, aes(sample = Performance_log)) + geom_qq()  + facet_wrap(~Group) + theme_classic()

qq+ geom_qq_line()


#Normality Tests
describe(lab4data$Performance_log)
shapiro.test(lab4data$Performance_log)

#Use the describeBy() function to get skewness and kurtosis by group
?describeBy()
describeBy(Performance_log ~ Group, data = lab4data)

#Use the group by function to get shapiro test results by group

lab4data %>%
  group_by(Group) %>%
  summarize(W = shapiro.test(Performance_log)$statistic, p_value = shapiro.test(Performance_log)$p.value)

#Use the filter function to get both

lab4data_clean <- drop_na(lab4data)
var(lab4data_clean$Performance_log)

lab4data_clean %>%
  group_by(Group) %>%
  summarize(variance = var(Performance_log))


leveneTest(Performance_log~Group, lab4data)

MANOVA(lab4data, dv = "Performance_log", between = "Group")

#What if you want to test equal variance between 2 groups specifically? 

lab4dataConG1<-lab4data %>%
  filter(Group == "Control" | Group == "G1")

leveneTest(Performance_log~Group, lab4dataConG1)

```

# Q3: Run the above tests again with the transformed outcome. Compare the differences in results.
The distribution was normalized. 
